---
title: "Programming in R"
author: "Jayesh Thakur"
format: html
editor: visual
editor_options: 
  chunk_output_type: console
---

Webpage created for submission of HW2 ST-558

## Task 1: Basic Vector practice

1.  **Creating two vectors for pre and post treatment**

```{r}
pre_treatment <- c(120, 151, 125, 126, 115, 132, 132, 129, 134, 139, 127, 122, 127, 135, 133, 128, 147, 138, 140, 132)

post_treatment <- c(127, 145, 135, 122, 115, 122, 123, 126, 126, 129, 132, 146, 120, 114, 121, 120, 128, 120, 133, 115)

pre_treatment
post_treatment
```

2.  **Assigning name to each observation in both vectors**

```{r}
subject_names <- paste("Subject", 1:20, sep = "_")

names(pre_treatment) <- subject_names
names(post_treatment) <- subject_names
```

3.  **Calculating the change in blood pressure before and after treatment**

```{r}
change_in_bp <- pre_treatment - post_treatment
change_in_bp
```

4.  **Calculating average decrease in blood pressure using mean()**

```{r}
avg_change <- mean(change_in_bp)
avg_change
```

5.  **Indices of Patients who has decreased blood pressure**

```{r}
pos_ind <- which(change_in_bp > 0)
pos_ind

names(pos_ind)
```

6.  **Creating a Subset vector to get patient with positive change i.e., decreased blood pressure**

```{r}
pos_change <- change_in_bp[change_in_bp > 0]
pos_change
```

7.  **Calculating Average decrease in blood pressure using mean()**

```{r}
avg_pos_change <- mean(pos_change)
avg_pos_change
```

## Task 2: Basic Data Frame practice

1.  **Creating DataFrame with 4 columns**, patient, pre_bp, post_bp, and diff_bp

```{r}
bp_df <- data.frame(
  patient = subject_names,
  pre_bp = pre_treatment,
  post_bp = post_treatment,
  diff_bp = pre_treatment - post_treatment
)

bp_df
```

2.  **Creating subset where diff_bp is negative** using subset()

```{r}
negative_diff_alt <- subset(bp_df, diff_bp < 0)
negative_diff_alt
```

3.  **Creating new column 'post_bp_lt_120'** indicating post_bp is less than 120

```{r}
bp_df$post_bp_lt_120 <- bp_df$post_bp < 120

bp_df
```

4.  **Displaying DataFrame with knitr**

```{r}
knitr::kable(bp_df)
```

## Task 3: List Practice

1.  **Creating placebo treatment DataFrame** along with column which indicates blood pressure less than 120

```{r}
placebo_pre_treatment <- c(138, 135, 147, 117, 152, 134, 114, 121, 131, 130)
placebo_post_treatment <- c(105, 136, 123, 130, 134, 143, 135, 139, 120, 124)

placebo_subject_names <- paste("Subject", 1:10, sep = "_")

placebo_df <- data.frame(
  patient = placebo_subject_names,
  pre_bp = placebo_pre_treatment,
  post_bp = placebo_post_treatment,
  diff_bp = placebo_pre_treatment - placebo_post_treatment
)

placebo_df$post_bp_lt_120 <- placebo_df$post_bp < 120

placebo_df
```

2.  **Creating list with Actual and placebo data**

```{r}
experiment_list <- list(
  treatment = bp_df,
  placebo = placebo_df
)

experiment_list
```

3.  **Access 1st element** i.e., treatment data using 3 methods

    1.  Using \$

    ```{r}
    experiment_list$treatment
    ```

    2.  Using \[\[ : *This returns as DataFrame*

    ```{r}
    experiment_list[["treatment"]]
    ```

    3.  Using \[ : *This returns as a list*

    ```{r}
    experiment_list["treatment"]
    ```

4.  **Accessing 'pre_bp' column from placebo dataframe**

```{r}
experiment_list$placebo$pre_bp
```
